An Improved Method for Coupling Hydrodynamics with Astrophysical Reaction Networks

### 2021-12-29 simplified_sdc_intro/paper.tex
more updates

\documentclass[times,modern]{aastex63}
\documentclass[times,preprint]{aastex63}

@@

We now look at the evolution of the state while burning by storing the
thermodynamic state internal to VODE each substep as it evaluates the
righthand side of the ODE system it is integrating.  We pick a point
on the symmetry axis a few zones behind the inward moving shock.
Figure~\ref{fig:nonaka_plot} shows the evolution of \isot{Ni}{56} over
a single coarse-grid timestep (4 fine-grid timesteps).

>>>>

%\documentclass[times,modern]{aastex63}
\documentclass[times,preprint]{aastex63}

@@

We now look at the evolution of the state while burning by storing the
thermodynamic state internal to VODE each substep as it evaluates the
righthand side of the ODE system it is integrating.  We pick a point
on the symmetry axis a few zones behind the inward moving shock and
ran both a Strang and simplified-SDC simulation for a single
coarse-grid timestep.  To ensure that we start off the same, both
simulations were started from a checkpoint file from the
simplified-SDC simulation a few steps before 0.1~s.
Figure~\ref{fig:nonaka_plot} shows the evolution of \isot{Ni}{56} over
a single coarse-grid timestep (4 fine-grid timesteps).  We see that
the Strang solution is much more discontinuous, as the reactions and
advection push the solution in different ways, because the processes
are completely independent of one another.  If an equilibrium state is
important to the nucleosynthesis, the large deviations in the Strang
solution mean that we are evaluating the reaction network with a state
out of equilibrium for much of the integration time. \MarginPar{need to double check the data---why are there points in the vertical part of Strang?}

\begin{figure}
\plotone{sdc_compare}
\caption{\label{fig:nonaka_plot} Evolution of $X(\isotm{Ni}{56})$ over
  4 fine-grid timesteps in a single zone behind the inward propagating
  shock.  The Strang points (blue) show the first $\Delta t/2$ burn
  bring the abundance away from equilibrium, then the advection
  overcorrecting it, and finally the second $\Delta t/2$ bringing it
  back toward the equilibrium.  The two simplified-SDC iterations
  appear much smoother.  For the SDC runs, there are a few points
  evaluated beyond the stop time---VODE can use data from outside the
  integration range to interpolate to the final solution.}
\end{figure}
